---
title: "abalone"
author: "Yuting"
date: "2019年11月23日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Proposal
# 1. Inspect correlations
#     - sex length and diameter highest correlation, no death
# 1. Box plots
#     - comparative boxplots for the different sex categories
# 1. Outliers?
#     - hov?
#         - if not homogeneous, transform the data
# 1. fit a model
#     - rings (or age) ~ (nothing specified)
# 1. Derived variable --> mature = male + female vs infant
#     - boxPlot against length and diameter
# 1. Derived variable --> volume = length*diameter*height
# 1. Multiple regression: whole ~ shucked + viscera + shell
# 
# 
# # Week 2: Fulfilling the proposal, collaborating with git
# 
# You all had several good initial analysis ideas.
# Unfortunately, your programmer was fired.
# You have to code it yourself.
# 
# Let's break the problem down into the following components:
# 
# 1. Data cleaning
#     - Please removed the 2 outliers from the `height` variable and re-run the `pairs` plot.
# 1. Variable manipulation
#     - Create a new variable "mature" based on "sex"
#         - Male and female are mature, infants are not
#     - Create a new variable "volume" based on "length", "diameter", and "height"


## data loading and cleaning
```{r}
library(ggplot2)
library(ggthemes)
library(GGally)
library(caret)
```
```{r}
data <- read.table("abalone(1).data",sep = ",")
head(data)
names(data) <- c("sex","length","diameter", "height", "whole", "shucked", "viscera", "shell", "rings")
head(data)
#check data structure
str(data) 
#change rings variable from integer to numeric
data$rings <- as.numeric(data$rings) 
```
```{r}

# order data by height in ascending order
data <- data[order(data$height) , ]
# check for outliers using head and tail functions. get row# which has outliers for height
head(data)
tail(data)
# row 1 and 2 in ordered data has value 0 for height, we will remove those two rows and save data frame as data.1
data.1<- data[-c(1:2,4176:4177),]
tail(data.1)
```


##variable manipulation
#Create a new variable "mature" based on "sex"
```{r}
## make new dummy variable 'mature' for sex
data.1$mature<-as.character(data.1$sex)
# Create a dummy variable for immature (0) and mature, M and F assigned 1
data.1$mature <- ifelse(data.1$mature == "I", 0,1)
data.1$mature <- as.factor(data.1$mature)
#drop the sex feature (highly related with mature)
data.2 <- data.1[-c(1)]
```

##Create a new variable "volume" based on "length", "diameter", and "height"
```{r}
## create a new variable volume (formula = area * height * diameter)
data.1$volume <- data.1$height*data.1$length*data.1$diameter
#same for data.2
data.2$volume <- data.2$height*data.2$length*data.2$diameter
```

##visualization
#plot metrics by sex
```{r}
## visualize attributes by sex
library(gridExtra)
# pairs plot to visualize data by sex (3 categories)
sexplot<-ggpairs(data.1, aes(colour=sex, alpha=0.8), title ="Pairs plot for abalone dataset by sex")+
  theme_grey(base_size = 8)
# pairs plot to visualize data by maturity (2 categories)
matplot<-ggpairs(data.1, aes(colour=as.factor(mature), alpha=0.8), title ="Pairs plot for abalone dataset by maturity")+
  theme_grey(base_size = 8)
#create list to put graphs into to then merge into pdf
plots = list(sexplot,matplot)

#save file
pdf(file = "plot.pdf")
plots
dev.off()
```

#model fit
#Model fit: Can we predict shucked, viscera, and shell weight based on whole weight?
```{r}
# models to predict shucked, viscera, shell weights from whole weight
# check how accurately whole weight predicts shucked, viscera, shell weights
fit.1 <- lm(shucked ~ whole + mature + whole*mature, data=data.2 )
summary(fit.1)
fit.2 <- lm(viscera ~ whole*mature, data=data.2)
summary(fit.2)
fit.3 <- lm(shell ~ whole*mature, data = data.2)
summary(fit.3)
# looks like we can pretty accurately predict shucked, viscera and shell weights from whole weight.
```

#Model fit: Perform an initial model fit for the data.
```{r}
#lm
model.fit <- lm(rings ~ length + diameter + height + whole + mature , data=data.2)
summary(model.fit)
model.fit2 <- lm(rings ~ volume + whole + mature , data=data.2)
summary(model.fit2)
model.fit3 <- lm(rings ~ length + diameter + whole + mature , data=data.2)
summary(model.fit3)
anova(model.fit,model.fit2)

#try stepwise
library(tidyverse)
library(caret)
library(leaps)
library(MASS)
#fit the full model
full.model <- lm(rings~.,data = data.2)
#stepwise regression model
step.model <-stepAIC(full.model,direction = "both", trace = FALSE)
summary(step.model)

#cross validation
set.seed(42)
#set up reapted k-flod cross-validation
train.control <- trainControl(method = "cv", number = 10)
#train the model
step.model.lm <- train(rings~., data = data.2, 
                    method = "leapBackward",
                    tuneGrid = data.frame(nvmax = 1:9),
                    trControl = train.control
                    )
#get the results
step.model.lm$results
#best tune
step.model.lm$bestTune
#9
#summary
summary(step.model.lm$finalModel)
#coefficients
coef(step.model.lm$finalModel, 9)
```

#glm fit
```{r}
model.fit4 <- glm(rings ~ whole + mature + volume, data = data.2, family = poisson(link = "log"))
summary(model.fit4)
#0.64
model.fit5 <- glm(rings ~ whole + mature + volume, data = data.2, family = gaussian(link = "identity"))
summary(model.fit5)
#R^2 0.66
model.fit6 <- glm(rings ~ whole + mature + height, data = data.2, family = gaussian(link = "identity"))
summary(model.fit6)

#stepwise for glm
#fit the full glm model
full.model.glm <- glm(rings~.,family = gaussian, data = data.2)
#stepwise regression model
step.model.glm <- stepAIC(full.model.glm,direction = "both", trace = FALSE)
summary(step.model.glm)
```

## Simulation

Please simulate a set of 1250 new observations based on your best model.
How do these simulated observations compare to the real observations?

F-statistic:   576 on 9 and 4163 DF,  p-value: < 2.2e-16

```{r}
length <- seq(min(data.2$length),max(data.2$length),length.out = 5)
diameter <- seq(min(data.2$diameter),max(data.2$diameter),length.out = 5)
height <- seq(min(data.2$height),max(data.2$height),length.out = 5)
whole <- seq(min(data.2$whole),max(data.2$whole),length.out = 5)
mature <- c(0,1)

grid <- expand.grid(length,diameter,height,whole,mature)
names(grid) <- c("length","diameter","height","whole","mature")

set.seed(90210)
grid$error <- rnorm(n = 1250,mean = 0,sd = 2.513)

library(tidyverse)

grid_sim <- grid%>%
  mutate(y=3.4617-11.52*length+18.38*diameter+42*height-0.7*whole-0.91*mature+error)

grid_sim$mature <- as.factor(grid_sim$mature)
grid_sim$pred <- predict(model.fit,newdata=grid_sim)
head(grid_sim)
plot(x = grid_sim$y,y=grid_sim$pred)
abline(a = 0,b = 1)

grid$error <- rnorm(1250,0,2.513)
head(grid)
library(tidyverse)

grid_sim <- grid%>%
  mutate(y=3.4617-11.52*length+18.38*diameter+42*height-0.7*whole-0.91*mature+error)
grid_sim$mature <- as.factor(grid_sim$mature)
grid_sim$pred <- predict(model.fit,newdata=grid_sim)
head(grid_sim)
plot(x = grid_sim$y,y=grid_sim$pred)
abline(a = 0,b = 1)
```


